---
title: "Generating Exams and Exercises with extexams"
author: "Joao Batista Tolentino Jr."
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Getting Started with extexams}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)

library(extexams)
```

## Introduction

The `extexams` package streamlines the creation of dynamic academic assessments. Built on top of the robust `exams` package, it provides tailored functions for two primary teaching workflows: generating batch exercises for weekly practice, and generating personalized, roster-based exams.


---

## 1. Batch Exercises (`generate_exercises`)

For regular assignments or weekly practice, `generate_exercises()` creates multiple randomized copies of a set of questions. By tying the random seed to the `semester` string, you ensure that the generated documents remain reproducible.

```{r batch-exercises, eval=FALSE}
# Example: Generating 3 copies of a practice sheet covering ANOVA and Fruit characteristics

# We use a temporary directory for the output in this example to comply with CRAN policies.
temp_dir <- tempdir()

exercise_report <- generate_exercises(
  exercises = c("anova", "fruit"),
  course = "Estatistica",
  semester = "2026.1",
  n_exerc = 3,
  output_dir = temp_dir
)

# View the metadata and answer keys for the generated batch
head(exercise_report)
```

---

## 2. Personalized Roster Exams (`generate_exams`)

When administering high-stakes evaluations, `generate_exams()` allows you to map a list of exercises directly to a student roster (`data.frame`). 

A unique feature of this function is the `substitution_list`. If a student has already demonstrated mastery of a specific topic (e.g., scoring > 0 in a previous cycle), the function can automatically swap that specific exercise for a "feito" (done) placeholder, saving them from redundant testing.

```{r roster-exams, eval=FALSE}
# 1. Create a mock student roster
students <- data.frame(
  Nome = c("Joao", "Maria", "Carlos"),
  P1_ciclo = c(1, 0, 0) # Joao passed the first cycle, Maria and Carlos did not
)

# 2. Define the exam structure
exs <- list("anova", "fruit")

# 3. Generate the exams
exam_keys <- generate_exams(
  exercises = exs,
  course = "Estatistica",
  semester = "2026.1",
  exam_id = "P1",
  data = students,
  name_col = "Nome",
  substitution_list = list(P1_ciclo = 1), # Replaces the 1st exercise ("anova") if P1_ciclo > 0
  output_dir = temp_dir
)

# View the concatenated answer keys for all students
head(exam_keys)
```

In the example above, Joao will receive a PDF where the ANOVA question is replaced by the `feito.Rmd` template, while Maria and Carlos will receive the full exam.


---
